

Read in the VB Player matches Data, NOTE we are already in the correct repository


```{r q1sol}
setwd("/home/jovyan/Mentorship_with_Michael_Code/Data")
vb <- read.csv("vb_player_matches.csv")
```


Okay great, the data is in. Let's look at row 1 of *vb* 


```{r q2sol}
vb[1,]
```



Let's look at column 1 of *vb* 


```{r q3sol}
vb[,1]
```



Let's look at rows 1,2,3,4, and 20 of *vb* 


```{r q4sol}
vb[c(1,2,3,4,20),]
#or
vb[c(1:4,20),]
```



Let's look at rows 1,2,3 and columns 2,3,4 of *vb* 


```{r q5sol}
vb[c(1,2,3),c(2,3,4)]
#or
vb[1:3,2:4]
```


Let's look at rows 1,2,3 and columns 2,3,4 of *vb* 


```{r q6sol}
vb[c(1,2,3),c(2,3,4)]
#or
vb[1:3,2:4]
```


Make a birthday column!

Step 1: Convert date to a date variable by running this code `vb$date <- as.Date(vb$Date)`
Step 2: Round each player's age by doing `vb$age <- round(vb$age)`
Step 3 (you finish it off): Create a new column called `bday` You can use base R or dplyr

Hint: the function `years()` placed aroud `age` should help for date subtraction :) 



```{r q7sol}
vb$date <- as.Date(vb$date)
vb$age <- round(vb$age)


#Base R solution

vb$bday <- vb$date-years(vb$age)

#Mutate

library(dplyr)

vb <- vb%>%
  mutate(bday=date-years(age))
```


Dplyr --> `arrange` vb by year and then descending `country` names.


```{r q8sol}
vb%>%
  arrange(year,desc(country))
```


Dplyr --> `arrange` vb by year and then descending `country` names and then `select` the `name` , `age` and `hgt` columns... assign this to an object called `metrics`


```{r q9sol}
metrics <- vb%>%
  arrange(year,desc(country))%>%
  select(name,age,hgt)
```


Read in `performance_joined.csv` ... assign it to an object named `perform`

```{r q10sol}
setwd("/home/jovyan/Mentorship_with_Michael_Code/Data")
perform <- read.csv("performance_joined.csv")
```


Read in the `skimr` library. `skim()` the `perform` object

```{r q11sol}
library(skimr)
skim(perform)
```


Using the `perform` dataframe. Create a new data frame named `us_names` where you

(1) Filter for players from the United states and who have over 1000 `n_matches`
(2) Select the name column
(3) Arrange by name


this should create an alphabetized name list from the US with players who had over 1000 games

save this as a csv named, `alpha_us.csv`

```{r q12sol}
write.csv(
  perform%>%
  filter(country=="United States",n_matches>1000)%>%
  select(name)%>%
  arrange(name), "alpha_us.csv"
)
```

Let's visualize the `perform` data frame. Start by viewing some densities

Using ggplot

Let's look at the density for `avg_errors.` 

Make sure to fill the density by `gender` set alpha to 0.4

make the theme black and white.

```{r q13sol}
perform%>%
  ggplot(aes(x=avg_errors))+
  geom_density(aes(fill=gender),alpha=0.4)+
  theme_bw()
```


Do the same thing as above, but with a histogram.

This time add a title using `+ggtitle` ... call it "test title"  

```{r q14sol}
perform%>%
  ggplot(aes(x=avg_errors))+
  geom_histogram(aes(fill=gender),alpha=0.4)+
  theme_bw()
```


Do the same thing as above, but with a histogram.


but now save this a pdf called "test.pdf"

```{r q15sol}
pdf("test.pdf")

perform%>%
  ggplot(aes(x=avg_errors))+
  geom_histogram(aes(fill=gender),alpha=0.4)+
  theme_bw()

dev.off()
```


First filter where `n_matches >500`

Let's look at two variables at once. Plot `avg_attacks` (y axis) vs. `avg_kills` (x axis)

Use a scatter plot with geom_point
color the points locally by gender
facet by gender
use a black and white theme
fit a the following line `+geom_smooth(method="lm")`

```{r q16sol}
perform%>%
  filter(n_matches>200)%>%
  ggplot(aes(x=avg_kills,y=avg_attacks))+
  geom_point(aes(color=gender))+
  facet_wrap(~gender)+
  theme_bw()+
  geom_smooth(method="lm")

```


Call the plotly library and put `ggplotly()` around the entire last graphic

```{r q17sol}
library(ggplotly)

ggplotly(perform%>%
  filter(n_matches>200)%>%
  ggplot(aes(x=avg_kills,y=avg_attacks))+
  geom_point(aes(color=gender))+
  facet_wrap(~gender)+
  theme_bw()+
  geom_smooth(method="lm"))

```


