

Read in the VB Player matches Data, NOTE we are already in the correct repository


```{r q1}

```


Okay great, the data is in. Let's look at row 1 of *vb* 


```{r q2}


```



Let's look at column 1 of *vb* 

```{r q3}

```



Let's look at rows 1,2,3,4, and 20 of *vb* 


```{r q4}


```



Let's look at rows 1,2,3 and columns 2,3,4 of *vb* 


```{r q5}

```


Let's look at rows 1,2,3 and columns 2,3,4 of *vb* 


```{r q6}

```


Make a birthday column!

Step 1: Convert date to a date variable by running this code `vb$date <- as.Date(vb$Date)`
Step 2: Round each player's age by doing `vb$age <- round(vb$age)`
Step 3 (you finish it off): Create a new column called `bday` You can use base R or dplyr

Hint: the function `years()` placed aroud `age` should help for date subtraction :) 



```{r q7}


```


Dplyr --> `arrange` vb by year and then descending `country` names.


```{r q8}

```


Dplyr --> `arrange` vb by year and then descending `country` names and then `select` the `name` , `age` and `hgt` columns... assign this to an object called `metrics`


```{r q9}

```


Read in `performance_joined.csv` ... assign it to an object named `perform`

```{r q10}

```


Read in the `skimr` library. `skim()` the `perform` object

```{r q11}

```


Using the `perform` dataframe. Create a new data frame named `us_names` where you

(1) Filter for players from the United states and who have over 1000 `n_matches`
(2) Select the name column
(3) Arrange by name


this should create an alphabetized name list from the US with players who had over 1000 games

save this as a csv named, `alpha_us.csv`

```{r q12}

```

Let's visualize the `perform` data frame. Start by viewing some densities

Using ggplot

Let's look at the density for `avg_errors.` 

Make sure to fill the density by `gender` set alpha to 0.4

make the theme black and white.

```{r q13}

```


Do the same thing as above, but with a histogram.

This time add a title using `+ggtitle` ... call it "test title"  

```{r q14}

```


Do the same thing as above, but with a histogram.


but now save this a pdf called "test.pdf"

```{r q15}

```


First filter where `n_matches >500`

Let's look at two variables at once. Plot `avg_attacks` (y axis) vs. `avg_kills` (x axis)

Use a scatter plot with geom_point
color the points locally by gender
facet by gender
use a black and white theme
fit a the following line `+geom_smooth(method="lm")`

```{r q16}


```


Call the plotly library and put `ggplotly()` around the entire last graphic

```{r q17}


```


