  
--- 
title: Assignment 1:Intro to Stat I!"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn the basics of R. Learn how to navigate the interactive `learnr` lessons.
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(nycflights13)
library(Lahman)
tutorial_options(exercise.timelimit = 60)
knitr::opts_chunk$set(error = TRUE)
```

## Tuesday 1 Assignment!


```{r quiz1, echo=FALSE}
quiz(caption = "Assignment 1",
    question("Which of the following would NOT be output from sample(1:4, 3, replace=FALSE)",
    answer("1,2,4"),
    answer("1,1,4", correct = TRUE),
    answer("2,3,4"),
    answer("1,3,4"),
    correct = "We are choosing from 1,2,3,4. Once a number is selected it CANNOT be selected again since replace=FALSE",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ),
  question("Consider the following statements. What are the answers: (i) 10 > 10     (ii) 14  <= 14    (iii) 100!=100    (iv) (10<12)&&(20!=20) ",
    answer("FALSE, TRUE, FALSE, FALSE",correct=TRUE),
    answer("FALSE, FALSE, FALSE, FALSE"),
    answer("FALSE, TRUE, FALSE, TRUE"),
    answer("FALSE, TRUE, TRUE, FALSE"),
    correct = "Refer to Slide 11",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ),
   question("Which function can we use to identify loaded packages?",
    answer("search()",correct=TRUE),
    answer("seek()"),
    answer("finder()"),
   correct = "Refer to Slide 15",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ),
  question("Suppose I'm working with the data frame next_gen. Suppose Mahomes' 2019 stats are in Row 10 and Passing Yards (listed as p_yds) are in column 5. Select all of the following that result in Mahomes' 2019 passing yards?" ,
    answer("next_gen[10,5]",correct=TRUE),
    answer("next_gen[5,10]"),
    answer("next_gen$p_yds[10]",correct=TRUE),
    answer("next_gen$p_yds[5]"),
   correct = "Refer to Slide 18",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ),
   question("When plotting variables using the baseline plot() function, how do we label the x-axis?" ,
    answer("xlab=SOMETHING IN QUOTES",correct=TRUE),
    answer("xlabel=SOMETHING IN QUOTES"),
    answer("xlabels=SOMETHING IN QUOTES"),
    answer("x=SOMETHING IN QUOTES"),
   correct = "Refer to Slide 23",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ),
  question("Which variable is categorical/qualitative?" ,
    answer("QB's cell phone number",correct=TRUE),
    answer("QB's height"),
    answer("QB's 40-yd dash time"),
    answer("QB's longest TD pass of the season"),
   correct = "Refer to Slide 28",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ),
  
   question("The GM of your team is interested in how your QB's passing yards per/plays where the QB leaves the pocket compares to other QBs (when they leave the pocket). Which visualization is most appropriate?" ,
    answer("A histogram of all QB passing yds/play (given they left the pocket), with a vertical line showing your QB's data.",correct=TRUE),
    answer("A bar plot of all QB passing yds/play (given they left the pocket), with a vertical line showing your QB's data."),
    answer("A scatter plot of all QB passing yds/play (given they left the pocket), with a vertical line showing your QB's data."),
   correct = "Refer to Slide 32",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ),
  
   question("The area under ANY DENSITY" ,
    answer("Is equal to 1",correct=TRUE),
    answer("Is equal to 100"),
    answer("Changes depending on the data"),
   correct = "Refer to Slide 33",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ),
  
 question("Which of the following are ROBUST to outliers?" ,
    answer("Median",correct=TRUE),
    answer("IQR",correct=TRUE),
    answer("Mean"),
    answer("Standard Deviation"),
   correct = "Refer to Slides 39-41",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ),
 
 question("Which of the following datasets will have the SMALLEST variance?" ,
    answer("4,4,4,100"),
    answer("4,4,4.1,4.1",correct=TRUE),
    answer("4,8,12,16"),
   correct = "Refer to Slide 46",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ),
  
   question("Suppose NBA players have heights that are normally distributed with a mean of 74 inches and standard deviation of 3 inches. 77 inches represents which percentile?" ,
    answer("50"),
    answer("84",correct=TRUE),
    answer("16"),
   correct = "Refer to Slides 48-58",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ),
  
    question("What is the z-score in the previous question" ,
    answer("-1"),
    answer("1",correct=TRUE),
    answer("2"),
    answer("-2"),
   correct = "Refer to Slides 48-58",
    random_answer_order = TRUE,
    allow_retry = TRUE
  )
  
)
```


### Coding Exercises

1.  Appropriately label the graphic (x-axis, y-axis, and title)

    ```{r submodex16, exercise = TRUE}
  #Data
next_gen <- suppressMessages(read_csv("https://raw.githubusercontent.com/ezczstat/bootcamp2020/master/next_gen_season.csv"))

#Enter Answer Below
    plot(next_gen$completionPercentage,next_gen$passerRating,
     xlab="_______",
     ylab="_______",
     main="________"
     )

    ```
    ```{r submodex16-solution}
 
next_gen <- suppressMessages(read_csv("https://raw.githubusercontent.com/ezczstat/bootcamp2020/master/next_gen_season.csv"))


plot(next_gen$completionPercentage,next_gen$passerRating,
     xlab="Completion Percentage",
     ylab="Passer Rating",
     main="Passer Rating vs. Completion Percentage"
     )
    ```
    
    <div id="submodex16-hint">
    **Hint:** `You can customize spacing`
     <div id="submodex16-solution">
    **Solution (click the lightbulb above):**
    </div>


2.  Add a vertical line to the histogram for the mean Passer Rating. Make the line blue with a width of 3.

    ```{r submodex17, exercise = TRUE}
  #Data
next_gen <- suppressMessages(read_csv("https://raw.githubusercontent.com/ezczstat/bootcamp2020/master/next_gen_season.csv"))

#Enter Answer Below
    hist(next_gen$passerRating,
main="Passer Ratings since 2017",
xlab="Passer Rating",
col="red",
las=1) #las=1 rotate y values
"_______________"
 
    ```
    ```{r submodex17-solution}

next_gen <- suppressMessages(read_csv("https://raw.githubusercontent.com/ezczstat/bootcamp2020/master/next_gen_season.csv"))


  hist(next_gen$passerRating,
main="Passer Ratings since 2017",
xlab="Passer Rating",
col="red",
las=1) #las=1 rotate y values
abline(v=mean(next_gen$passerRating),
       col="blue",
       lwd=3)

    ```
    
    <div id="submodex17-hint">
    **Hint:** `SLIDE 50`
     <div id="submodex17-solution">
    **Solution (click the lightbulb above):**
    </div>
    
    
    
3. The plot below is from slide 48, can you adjust the histogram to show 24 bars (equal in width?)

    ```{r submodex18, exercise = TRUE}
  #Data
next_gen <- suppressMessages(read_csv("https://raw.githubusercontent.com/ezczstat/bootcamp2020/master/next_gen_season.csv"))

#Enter Answer Below
 hist(next_gen$passYards,
xlab="Passing Yards",
main="Passing Yards")
abline(v=mean(next_gen$passYards),
lwd=3,
col="red")
abline(v=c(mean(next_gen$passYards)
+sd(next_gen$passYards),
mean(next_gen$passYards)-
sd(next_gen$passYards)),
lwd=3,
col="blue")
"_______________"
 
    ```
    ```{r submodex18-solution}

next_gen <- suppressMessages(read_csv("https://raw.githubusercontent.com/ezczstat/bootcamp2020/master/next_gen_season.csv"))


 hist(next_gen$passYards,
xlab="Passing Yards",breaks=24,
main="Passing Yards")
abline(v=mean(next_gen$passYards),
lwd=3,
col="red")
abline(v=c(mean(next_gen$passYards)
+sd(next_gen$passYards),
mean(next_gen$passYards)-
sd(next_gen$passYards)),
lwd=3,
col="blue")

    ```
    
    <div id="submodex18-hint">
    **Hint:** `SLIDE 50`
     <div id="submodex18-solution">
    **Solution (click the lightbulb above):**
    </div>


4. Here are two histograms, avgTimeToThrow and aggressiveness. Enter a single line of code to output the graphics on top of one another.

    ```{r submodex19, exercise = TRUE}
  #Data
next_gen <- suppressMessages(read_csv("https://raw.githubusercontent.com/ezczstat/bootcamp2020/master/next_gen_season.csv"))

#Enter Answer Below
#THE ANSWER SHOULD APPEAR HERE, BEFORE THE FIRST HISTOGRAM ARGUMENT!
hist(next_gen$avgTimeToThrow,
#histogram
xlab="Average Time to Throw (sec)",#xaxis
main="Average Time to Throw", #title
col="red") #color
abline(v=mean(next_gen$avgTimeToThrow)
#vertical bar
,lwd=2) #line width
hist(next_gen$aggressiveness,
xlab="Aggressiveness %",
main="Aggressiveness",
col="blue")
abline(v=mean(next_gen$aggressiveness)
,lwd=2)
    ```
    ```{r submodex19-solution}
next_gen <- suppressMessages(read_csv("https://raw.githubusercontent.com/ezczstat/bootcamp2020/master/next_gen_season.csv"))


par(mfrow=c(2,1))
hist(next_gen$avgTimeToThrow,
#histogram
xlab="Average Time to Throw (sec)",#xaxis
main="Average Time to Throw", #title
col="red") #color
abline(v=mean(next_gen$avgTimeToThrow)
#vertical bar
,lwd=2) #line width
hist(next_gen$aggressiveness,
xlab="Aggressiveness %",
main="Aggressiveness",
col="blue")
abline(v=mean(next_gen$aggressiveness)
,lwd=2)

    ```
    
    <div id="submodex19-hint">
    **Hint:** `SLIDE 41`
     <div id="submodex19-solution">
    **Solution (click the lightbulb above):**
    </div>
    
    
    
5. Create a data frame named `best_rushers` with the following information: Variable 1 --> `players`:  Smith,  Payton,  Gore,  Sanders, and  Peterson. Variable 2 --> `rush_yds` --> 18355, 16726, 15347, 15269, 14216. Create a bar chart to compare the career rushing yards for each player. Make sure the bar plot is horizontal. Assign the following colors to each player Peterson = yellow, Sanders = green , Gore=Orange, Payton=blue, and Smith=Red.

    ```{r submodex20, exercise = TRUE}
  #Enter answer here
    
    ```
    ```{r submodex20-solution}
best_rushers <- data.frame(players=c("Smith","Payton","Gore","Sanders","Peterson"),
                           rush_yds=c(18355,16726,15347,15269,14216))


    barplot(best_rushers$rush_yds, 
            col=c("red","blue","orange","green","yellow"),
            names.arg = best_rushers$players,
            horiz=T,
            las=1)
    ```
    
    <div id="submodex20-hint">
    **Hint:** `to change to horizontal, horiz=T. To name the bars use names.arg=`
     <div id="submodex20-solution">
    **Solution (click the lightbulb above):**
    </div>